### 2016-02-06 paper/paper.tex
Add condensation types subsubsubsection

\RequirePackage{amsmath}
\documentclass{llncs}

\usepackage{amssymb}

\usepackage{cmap}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}

\usepackage{tikz}

\usepackage[font=small]{caption}
\usepackage{subcaption}

\usepackage{hyperref}
\usepackage{cleveref}

\usepackage{listings}
\usepackage{xcolor}

\usepackage{xspace}

\tikzset{every picture/.append style={scale=0.5}}

\tikzstyle{cell}=[rectangle,draw=gray,dashed,semithick, minimum size=1.05cm]
\tikzstyle{graybox}=[rectangle,thick, minimum size=0.4cm, fill=lightgray]
\tikzstyle{whitebox}=[rectangle,draw=black,thick, minimum size=0.5cm]

% \captionsetup[subfigure]{subrefformat=simple,labelformat=simple}
% \renewcommand\thesubfigure{(\alph{subfigure})}

\lstset{
  language=c++,
  basicstyle=\ttfamily \scriptsize,
  lineskip={-1.5pt},
  columns=fixed,
  basewidth=0.5em,
  keywordstyle=\bfseries\color{blue!60!black},
  commentstyle=\itshape\color{gray!80!black},
  identifierstyle=\color{violet!60!black},
  stringstyle=\color{orange},
  frame=single
  % numbers=left
}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand*{\N}{\mathbb{N} \xspace}
\DeclareMathOperator*{\argmin}{arg\,min}

\renewcommand{\C}{\texttt{C} \xspace}

\begin{document}

\title{Linear variation and an optimization of algorithms for connected
components labeling in a binary image}

\author{Fedor Alekseev\inst{1} \and Mikhail Alekseev\inst{2}\inst{3}
\and Artyom Makovetskii\inst{2}\inst{4}}

\institute{Moscow Institute of Physics and Technology State University, Dolgoprudny, Russia\\
\email{alekseev@phystech.edu}
\and
Chelyabinsk State University, Chelyabinsk, Russia
\and
\email{alexeev@csu.ru}
\and
\email{artemmac@mail.ru}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Linear variation is a topological characteristic of a function of two variables.
The problem of linear variation computing can be reduced to the problem
of counting connected components in a binary image with eight-connected connectivity.
The new hybrid method for the problem is presented.
The method is essentially a modification of a two-scan algorithm for labeling
that groups the pixels into $2 \times 2$ cells.

\keywords{Binary raster image, connected component labeling,
pattern recognition}
\end{abstract}

\section{Introduction}

% Makovetskii part

In image restoration it is often necessary to consider the following problem.
How to restore an original undistorted image $v$, if it is known an observed
image $u_0$ and the relation between $u_0$ and $v$:
\begin{equation}
  u_0=v+n,
  \label{eq:imageRelation}
\end{equation}
where $n$ is a noise?
A common way for solving the problems \eqref{eq:imageRelation} is to use the
variational functionals.
One of the most widely used approaches is total variation
regularization [1]. % TODO fix bibliography
Let $J(u)$ be the following functional:
\begin{equation}
  J(u) = \frac12 \norm{u - u_0}^2 + \lambda TV(u),
  \label{eq:jFunctional}
\end{equation}
where $\norm{\cdot}$ is the $L_2$ norm and all function we consider belong to the class
$BV(\Omega)$, i.\,e. the class of bounded variation on the set $\Omega$ functions.
%
The expression $\frac12 \norm{n-n_0}^2$ is called a fidelity term and
$\lambda TV(u)$ is called a regularization term and $\lambda$ is regularization
parameter.

\begin{equation}
  TV(u) = \int\limits_{\Omega} |\nabla u|\,dx\,dy.
  \label{eq:TV}
\end{equation}

Let $u_*$ be extremal function for the following variational problem:
\begin{equation}
  u_* = \argmin_{u \in BV(\Omega)} J(u).
  \label{eq:ustar}
\end{equation}
The \autoref{eq:ustar} is called total variation regularization problem.

The norm and modulus of the gradient are metrical characteristic of a function of
two variables.
Continuous functions of two variables also have a set of topological
characteristics referred to as linear variations.
Kronrod [2] % TODO fix bibliography
introduced the notion of a regular component of the level set of a
continuous function of two variables.
The simplest topological characteristic in the linear variation theory is a number
of regular components for all level sets of a function.
Full information about linear variations of a function is contained in the
one-dimensional tree of a function of two variables.
In [3, 4, 5] % TODO fix bibliography
was discussed the necessity of the using a linear variation in the
image restoration theory.

Let $\Phi_u(t)$ be the number of regular components of a level set $t$
for a continuous function.
The first Kronrod's linear variation is defined as
\begin{equation}
  V(u) = \int\limits_{-\infty}^{+\infty} \Phi_u(t)\,dt.
  \label{eq:KronrodVariation}
\end{equation}

Let $w$ be a binary discrete function $w=(w_{i,j})$, where $w_{i,j} \in \{0,1\}$,
for all pairs $(i,j)$.
A subset of such pairs $(i, j)$ when $w_{i,j}=1$ and all elements of the subset are
connected by the 8-connectivity, is called the connected component of the binary
function $w$.
For a number $k \in \N$ and a discrete function $u$ we define the following
indicator function $\chi$:
\begin{equation}
  \chi_k (u_{i,j}) =
    \begin{cases}
      1, & u_{i,j} \ge k \\
      0, & u_{i,j} < k
    \end{cases}.
  \label{eq:chiIndicator}
\end{equation}

\begin{definition}
  The number $V_k(u_{i,j})$ of connected components for a level $k$, $k \in \N$ of
  the discrete function $u$ is called the number of connected components of the 
  binary discrete function $\chi_k(u_{i,j})$.
\end{definition}

\begin{definition}
  The linear variation $V(u_{i,j})$ of a discrete function  is defined as follows:
  \begin{equation}
    V(u_{i,j}) = \sum_{k=0}^{+\infty} V_k(u_{i,j})
    \label{eq:V}
  \end{equation}
\end{definition}

Let us compute the discrete gradient $\nabla u_{i,j}$ of $u$ at $(i,j)$ as
\begin{equation}
  \nabla u_{i,j} = (u_{i+1,j} - u_{i,j}, u_{i,j+1} - u_{i,j}).
  \label{eq:gradient}
\end{equation}

Suppose that if the pair $(i,j)$ is outside of the domain of the function $u$, then 
$u_{i,j} = 0$.
The main problem of the computer implementation of the linear variation is 
computation of $V_k(u_{i,j})$ for given k.

% Alekseev part

This problem is equivalent to the problem of counting the number of connected
components (CC) in a binary image with 8-connectivity.
This is usually done through CC labeling, and many approaches for that are
known\cite{sterzh}.

This paper presents an efficient way to convert the initial binary image
with 8-connectivity into
a condensed (non-binary) image with new connectivity that is $2$ times smaller in
each dimension than the initial image.
Most approaches for CC labeling are still valid for the new image.
As time and memory consumptions of conventional approaches usually depend linearly on
the number of pixels, running them on the condensed image can be
more efficient than running on the initial image.

It is worth noting that our method exploits an important property of 8-connectivity
that does not hold for 4-connectivity, so it is not applicable for the latter case.

\section{$2\times2$ condensation}

We will use the fact that for the case of 8-connectivity all $4$ pixels of any
$2 \times 2$ square are adjacent to each other, and so are contained in the same CC
and should have same labels upon completion of the labeling algorithm.
So instead of labeling single pixels we can label $2 \times 2$ groups of pixels,
or ``cells''.

Consider a $2N \times 2M$ binary image\footnote{
For convenience in this paper we consider only input images with even size
in both dimensions. If this is not the case, one can easily append a row or
a column of background pixels to the image.}
specified by a binary predicate $b(x, y)$,
returning $1$ in case the pixel at the given coordinates is black (object pixel),
and $0$ otherwise (background pixel).
For convenience we use 0-indexation throughout the paper, so the topmost row has
row index $y=0$, and the leftmost column has column index $x=0$.

The condensed image will be of size $N \times M$.
Each pixel of the condensed image will unambiguously represent the configuration
of the corresponding $2\times 2$ cell of pixels of the initial binary image.
So the condensed image will contain pixels of $2^4 = 16$ colors.
We suggest the function $c(x, y)$ denoting the color of a pixel of condensed image
to be the following:
\begin{align*}
  c(x, y) &= 2^0 b(2x, 2y)   + 2^1 b(2x+1, 2y) \\
          &+ {} 2^2 b(2x, 2y+1) + 2^3 b(2x+1,2y+1)
\end{align*}

So the value of each pixel of the initial image is stored in the corresponding
bit in color of the corresponding pixel of the condensed image.
Note that if $c(x, y) = 0$ for some $(x, y)$, then there are no object pixels in
this cell, and no label should be assigned to this whole cell. The enumeration of
pixels inside one cell is shown on \autoref{fig:pixelsEnumeration}.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}
    \node[cell] at (0.5, 0.5) {};
    \foreach \x/\y [count=\i from 0] in {0/1,1/1,0/0,1/0} {
      \node[whitebox] at (\x,\y) {\i};
    }
  \end{tikzpicture}
  \caption{Cell pixels enumeration}
  \label{fig:pixelsEnumeration}
\end{figure}

We need also to define connectivity function for the condensed image.
We cannot actually just use 8-connectivity, as whether two cells are adjacent
cells of one CC depends not only on their coordinates, but also on their
configuration. See \autoref{fig:connectivity:examples} for examples.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 1/1, 3/0, 3/1}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Nearby cells that are not directly connected.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 1/1, 2/0, 3/1}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Directly connected cells.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,4.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 0/1, 1/1, 4/0, 4/1, 5/0}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Cells that are not nearby in terms of 8-connectivity, cannot be
    directly connected.}
  \end{subfigure}
  \caption{Cells connectivity examples}
  \label{fig:connectivity:examples}.
\end{figure}

However, for every variant of relative placement of two nearby cells there is
a mask of pixels of each cells that are important for direct connectivity of
this pair of cells.
So two cells are considered directly connected, if they are nearby
in terms of 8-connectivity and each of them contains at least one object pixel
in the mask of important pixels corresponding to relative placement of the cells.
The masks are shown in \autoref{fig:connectivity:masks}.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (0.5, 0.5) {};
      \node[graybox] at (0, 2) {2};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (0, 1) {0};
      \node[graybox] at (1, 1) {1};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 2.5) {};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 2) {2};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.25\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 0) {3};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 0) {2};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \caption{Masks of important pixels for all four cases of relative placement
  of two nearby cells. Those pixels are filled in gray, with their numbers within
  cell specified.}
  \label{fig:connectivity:masks}
\end{figure}

The idea is that if we naturally encode mask pixels as 4-bit numbers in the
way similar to how we defined $c$, we can express the predicate denoting whether
two nearby cells
are directly connected efficiently with just two bitwise and one logical AND
operations.
As an example, \autoref{fig:connectivity:code} compares possible codes in \C checking
if the labels of the current pixel or cell and the one right on top of it
should be same for classic 8-connectivity and for the new connectivity.
Note that as bitwise operations are usually relatively cheap, we added only a
little complication compared to decreasing the number of pixels to process by the 
factor of 4.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.475\linewidth}
    \centering
    \begin{lstlisting}
    if (b(x, y)) {
      if (b(x, y-1)) {
        // copy labels, or union label sets
      }
      // consider other directions
    }
    \end{lstlisting}
    \caption{Binary image with 8-connectivity.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.475\linewidth}
    \centering
    \begin{lstlisting}
    if (c(x, y)) {
      if ((c(x, y-1) & ((1<<2) | (1<<3)))
       && (c(x, y)   & ((1<<0) | (1<<1)))) {
        // copy labels, or union label sets
      }
      // consider other directions
    }
    \end{lstlisting}
    \caption{Condensed image with new connectivity.}
  \end{subfigure}
  \caption{Checking the nearby top pixel/cell.}
  \label{fig:connectivity:code}
\end{figure}

\section{Benchmarks}

In this section we present performance benchmarks of some popular methods for
counting CC both with and without the $2 \times 2$ cells optimization.

\subsection{Algorithms overview}

As each of these methods, omitting DTSUF could be used with both condensed and
non-condensed images, for convenience we denote height and width of the image
on input of each algorithm as $H$ and $W$.

Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
meaning the appropriate raster for non-condensed and condensed images.
When we mention neighbors of some pixel, we mean pixels that are directly connected
to it in terms of appropriate connectivity.

\subsubsection{DFS}

The first algorithm we chose for the benchmark is Depth First Search graph traversal.
Graph traversal algorithms are the general approach for CC labeling in graphs, and
DFS is arguably the most simple.
Although not intended for image labeling and
suboptimal in most cases, it still has linear in the number of pixels worst-case
complexity in both time and memory.

Like most approaches, the algorithm has an outer loop over all pixels of the image.
Each time an object pixel (non-empty cell, in case it is actually a cell)
which is not labeled yet is encountered, DFS traversal is launched from that pixel.
The traversal labels all the object pixels of the same CC, launching recursively
from all unlabeled neighbors of current pixel.

\subsubsection{SUF}

Another general approach for graph CC labeling which is more popular in image
labeling is using some kind of a data structure to maintain disjoint sets of pixels
that are considered to be in the same CC.

The structure is commonly called Union-Find, as the two essential operations
it is supposed to support are Union, which is used to merge some two of the
disjoint sets in one, and Find, which is used to find the representative element
of the set some element belongs to, so that we can always tell if two elements are
in the same set.
The structure can be implemented so that the amortized time complexity for each
operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
and $\alpha(HW) < 5$ for all remotely practical image sizes. % TODO links

The approach that is named SUF (Scan plus Union-Find) in this paper follows.

Pixels are processed in the natural order: from top to bottom and from
left to right, so each row is fully processed before all rows after it.
So for each object pixel we encounter there are at most 4 nearby pixels
that were already processed, see \autoref{fig:suf:neighbors}.

At first each object pixel is considered to belong to its own CC, so right before
considering the neighbors we increment the CC counter.
Moreover, it is associated with its own disjoint set in the Union-Find structure.
More specifically, object pixel at position $x, y$ is associated with
the set number $xW + y$ to guarantee absence of collisions.

The possible neighbors are processed in the order shown on \autoref{fig:suf:neighbors},
and for each neighbor that is object pixel, the Union operation is performed
on the two sets containing the current and the neighbor pixel.
In case the Union operation was successful, which means that the pixels
had belonged previously to different sets and we have just found a way to merge two
CC, the CC counter is decremented.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \foreach \x/\y [count=\i from 1] in {0/1,1/1,2/1,0/0} {
      \node[whitebox] at (\x,\y) {\i};
    }
    \node[graybox] at (1, 0) {};
    \node[whitebox] at (1, 0) {};
  \end{tikzpicture}
  \caption{Neighbor mask for SUF algorithm family.}
  \label{fig:suf:neighbors}
\end{figure}

Note that this algorithm outputs not the matrix of final labels, but only the number
of CC.
This is the place in which it differs from some common two-pass approaches for CC % TODO references
labeling, as it is an algorithm not for labeling, but just for counting CC.
The latter is exactly what is needed for the aforementioned linear variation problem.

Also we have to mention that for performance reasons the Union-Find internal structures
are allocated ahead to have the size of at least $HW$ to guarantee that
each object can be easily associated with unique starting set given only its
coordinates. So memory consumption of SUF is linear in the number of pixels.

\subsubsection{SUF2}

This approach is a modification of SUF designed to reduce memory consumption
and potential number of cache-misses generated by Union-Find operations.

The idea is to run through each row twice: first time to establish label equality
classes within row, and the second time to set unique equal label for all pixels of
same CC.
This gives us the opportunity to maintain just $O(W)$
different labels at each moment, and so the Union-Find structure can be allocated
for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.

The exact maximum number of label equality classes between two adjacent rows
depends on whether the given image is condensed.
For non-condensed images $W=2M$, and maximum number of label equality classes
in two adjacent rows is $M=\frac{W}2$, as two object pixels are both in adjacent
rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.

On the other hand, for condensed images $W=M$, and maximum number of label equality
class in two adjacent rows of cells is $2W=2M$.
See \autoref{fig:suf2:extremal} for illustration on the extremal cases.

\begin{figure}
  \centering
  \begin{subfigure}{0.475\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0,1,...,9} \foreach \y in {0,1} {
        \node[whitebox,dashed,semithick,draw=gray] at (\x,\y) {};
      }
      \foreach \x/\y in {0/0,2/1,4/1,6/0,8/1} {
        \node[whitebox] at (\x,\y) {};
      }
    \end{tikzpicture}
    \caption{Non-condensed image, up to 5 different labels on 10 columns.}
  \end{subfigure}
  \quad
  \begin{subfigure}{0.475\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5,...,8.5} \foreach \y in {0.5,2.5} {
        \node[cell] at (\x,\y) {};
      }
      \foreach \x in {0,2,...,8} \foreach \y in {0,3} {
        \node[whitebox] at (\x,\y) {};
      }
    \end{tikzpicture}
    \caption{Condensed image, up to 10 different labels on 5 columns.}
  \end{subfigure}
  \caption{Extremal cases for SUF2.}
  \label{fig:suf2:extremal}
\end{figure}

Our benchmarks show that, although SUF2 without condensation is generally more
efficient on noise images than SUF without condensation, the mentioned
consideration make the $2 \times 2$ optimization
of SUF2 less dramatic in means of performance increase factor.
Optimized SUF2 significantly outperforms optimized SUF only on
huge noise images with less than 60\% density, while falling behind on
small and medium-sized noise images.

Note that, while SUF2 is intended for CC counting, its memory
consumption is only $O(W)$, which is significant for embedded systems and
other resource-tight applications.

\subsubsection{DTSUF}

This modification of SUF was implemented to compare the optimization we present with
another optimization suggested by Wu, Otoo and Suzuki, % TODO reference
namely decision tree.
Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
applicable for condensed image, as it exploits similar observations on 8-connectivity.

As before, the goal is to reduce the number of pixel lookup and label copy operations.
This is achieved with careful consideration of only the necessary neighbors,
stopping once the set of one or two label operations that are needed is
determined. The decision tree for processing an object pixel is
shown on \autoref{fig:dtsuf}.

\begin{figure}
  \centering
  \begin{subfigure}{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \node[graybox] at (1,0) {};
      \foreach \x/\y/\a in {0/1/a,1/1/b,2/1/c,0/0/d,1/0/e} {
        \node[whitebox,font=\bf] at (\x,\y) {\a};
      }
    \end{tikzpicture}
    \caption{Neighbor enumeration.}
  \end{subfigure}
  \quad
  \begin{subfigure}{0.65\linewidth}
    \centering
    \begin{tikzpicture}[
      sibling distance=18em,
      lookup/.style = {shape=circle, draw, thick, align=center, font=\bf},
      action/.style = {shape=rectangle, rounded corners, draw, align=center, font=\scriptsize},
      result/.style = {font=\bf}
    ]
      \node[lookup] {b}
        child[level distance=3em,sibling distance=30em] {
          node[lookup] {c}
          child[level distance=5em, sibling distance=14em] {
            node[lookup] {a}
            child[level distance=6em, sibling distance=10em] {
              node[lookup] {d}
              child {
                node[action] {new label}
                edge from parent node[result,above] {$0$}
              }
              child {
                node[action] {union(e,d)}
                edge from parent node[result,above] {$1$}
              }
              edge from parent node[result,above] {$0$}
            }
            child {
              node[action] {union(e,a)}
              edge from parent node[result,above] {$1$}
            }
            edge from parent node[result,above] {$0$}
          }
          child[level distance=9em,sibling distance=24em] {
            node[lookup] {a}
            child[level distance= 6em, sibling distance=10em] {
              node[lookup] {d}
              child {
                node[action] {union(e,c)}
                edge from parent node[result,above] {$0$}
              }
              child {
                node[action] {union(e,c,d)}
                edge from parent node[result,above] {$1$}
              }
              edge from parent node[result,above] {$0$}
            }
            child[sibling distance=20em] {
              node[action] {union(e,c,a)}
              edge from parent node[result,above] {$1$}
            }
            edge from parent node[result,above] {$1$}
          }
          edge from parent node[result,above] {$0$}
        }
        child[level distance=2em] {
          node[action] {union(e,b)}
          edge from parent node[result,above] {$1$}
        };
    \end{tikzpicture}
    \caption{The decision tree.}
  \end{subfigure}
  \caption{DTSUF decision tree optimization.}
  \label{fig:dtsuf}
\end{figure}

\subsection{Condensation types}

For memory-tight applications such as embedded systems it can be useful to
utilise the suggested condensation in a lazy way, so that without actually
constructing the condensed image we use a function (``view'') calculating color of
each cell with 4 lookups to the original binary image.
This is probably most interesting combined with SUF2 algorithm, as it is a way
to win some time without increasing memory footprint.
However this lazy approach increases the number of image lookups.
So, according to our benchmark, this combination is efficient only for dense
($>30\%$) noise.

So the benchmark tables show three levels of optimization applied to
DFS, SUF and SUF2:
``no'' meaning no condensation, ``view'' meaning lazy memory-less condensation
and ``yes'' meaning that a temporary condensed image was constructed in memory.
For the latter case, the construction time was included into the overall measured
time of run.

\subsection{Results}

some results

\section{Summary and future work}

\newcommand{\inputtable}[2]{
  \begin{table}[p]
    \centering
    \input{../bench-out/#1x#2}
    \caption{Average milliseconds of cpu time per run on noise images $#1 \times #2$ px}
    \label{tab:noise#1x#2}
  \end{table} }

% \inputtable{200}{300}
% \inputtable{400}{600}
\inputtable{800}{1200}
% \inputtable{1200}{1800}
\inputtable{2000}{3000}
\inputtable{4000}{6000}

\begin{thebibliography}{5}

\bibitem {hechao}
He, L., Chao, Y., Suzuki, K., Wu, K.:
Fast connected-component labeling. Pattern Recognition. v42, Pages 1977-1987. (2009)

\bibitem {sterzh}
Стержанов, М.:
Методики выделения связных компонент в штриховых бинарных изображениях.
Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. 2010, с. 169-174.

\end{thebibliography}

\end{document}