### 2016-04-14 paper/errata.tex
Fix references

\documentclass{article}

\usepackage{cmap}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}

\usepackage{xr}
\externaldocument{paper}

\begin{document}

\title{Errata for Linear Variation and an Optimization of Algorithms for Connected
Components Labeling in a Binary Image}

\author{Fedor Alekseev, Mikhail Alekseev and Artyom Makovetskii}

\date{}

\maketitle

\begin{enumerate}
  \item The references page was revised: references on Kronrod and Fisher-Yates
    were removed, as well the extra number before one of the references on
    Makovetskii and Kober.
\end{enumerate}

\end{document}