### 2016-04-20 paper/paper.tex
Merge Mikhail's changes

\RequirePackage{amsmath}
\documentclass[hidelinks]{llncs}

\usepackage[russian,english]{babel}

\usepackage{amssymb}

\usepackage{cmap}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}

\usepackage{tikz}

\usepackage[font=small]{caption}
\usepackage{subcaption}

\usepackage{hyperref}
\usepackage{cleveref}

\usepackage{listings}
\usepackage{xcolor}

\usepackage{xspace}

\usepackage{tabularx}
\usepackage{graphicx}

\usepackage{verbatim}

\newcommand{\texten}[1]{#1}
\newcommand{\textru}[1]{}

%% Чтобы переключиться на русский язык в некоторых секциях, можно 
%% раскомментировать следующие две команды
% \renewcommand{\texten}[1]{}
% \renewcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}

%% Чтобы оставить оба языка, можно раскомментировать
%% следующие две команды
% \renewcommand{\texten}[1]{#1}
% \renewcommand{\textru}[1]{\\ \foreignlanguage{russian}{\emph{#1}}}

\tikzset{every picture/.append style={scale=0.5}}

\tikzstyle{cell}=[rectangle,draw=gray,dashed,semithick, minimum size=1.05cm]
\tikzstyle{graybox}=[rectangle,thick, minimum size=0.4cm, fill=lightgray]
\tikzstyle{whitebox}=[rectangle,draw=black,thick, minimum size=0.5cm]

% \captionsetup[subfigure]{subrefformat=simple,labelformat=simple}
% \renewcommand\thesubfigure{(\alph{subfigure})}

\lstset{
  language=c++,
  basicstyle=\ttfamily \scriptsize,
  lineskip={-1.5pt},
  columns=fixed,
  basewidth=0.5em,
  keywordstyle=\bfseries\color{blue!60!black},
  commentstyle=\itshape\color{gray!80!black},
  identifierstyle=\color{violet!60!black},
  stringstyle=\color{orange},
  frame=single
  % numbers=left
}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand*{\N}{\mathbb{N} \xspace}
\DeclareMathOperator*{\argmin}{arg\,min}

\renewcommand{\C}{\texttt{C} \xspace}
\newcommand{\CXX}{\texttt{C++} \xspace}

\begin{document}

\title{Linear Variation and an Optimization of Algorithms for Connected
Components Labeling in a Binary Image}

\author{Fedor Alekseev\inst{1} \and Mikhail Alekseev\inst{2}\inst{3}
\and Artyom Makovetskii\inst{2}\inst{4}}

\institute{Moscow Institute of Physics and Technology State University, Dolgoprudny, Russia\\
\email{alekseev@phystech.edu}
\and
Chelyabinsk State University, Chelyabinsk, Russia
\and
\email{alexeev@csu.ru}
\and
\email{artemmac@mail.ru}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The linear variation is a topological characteristic of the function of two
variables. The problem of linear variation computing can be reduced to the
problem of counting connected components in a binary image with eight-connected
connectivity.
%The new hybrid method for the problem is presented.
%The method is essentially a modification of a two-scan algorithm for labeling
%that groups the pixels into $2 \times 2$ cells.
%This is usually done through connected components labeling, and many approaches
%for that are known. This paper presents an efficient way to convert the initial
%binary image with 8-connectivity into a condensed %(non-binary) image with new
%connectivity that is 2 times smaller in each dimension than the initial image.
%Most approaches for connected components labeling are still valid for the new
%image. As time and memory consumptions of conventional approaches usually
%depend linearly on the number of pixels, running them on the condensed image
%can be up to 4 times more efficient than running on the initial image.
The method is essentially a modification of some known raster algorithms for
labeling that groups pixels into $2 \times 2$ cells. A performance benchmark of
the proposed  method applied to some approaches on noise images is provided.

\keywords{Binary raster image, connected component labeling, pattern
recognition}
\end{abstract}

\section{Introduction}

% Makovetskii part

\begin{comment}
In image restoration it is often necessary to consider the following problem.
How to restore an original undistorted image $v$, if it is known an observed
image $u_0$ and the relation between $u_0$ and $v$:
\begin{equation}
  u_0=v+n,
  \label{eq:imageRelation}
\end{equation}
where $n$ is a noise?
A common way for solving the problems \eqref{eq:imageRelation} is to use the
variational functionals.
One of the most widely used approaches is total variation
regularization\cite{mak1}.
Let $J(u)$ be the following functional:
\begin{equation}
  J(u) = \frac12 \norm{u - u_0}^2 + \lambda TV(u),
  \label{eq:jFunctional}
\end{equation}
where $\norm{\cdot}$ is the $L_2$ norm and all function we consider belong to
the class $BV(\Omega)$, i.\,e. the class of bounded variation on the set
$\Omega$ functions.
%
The expression $\frac12 \norm{n-n_0}^2$ is called a fidelity term and
$\lambda TV(u)$ is called a regularization term and $\lambda$ is regularization
parameter.

\begin{equation}
  TV(u) = \int\limits_{\Omega} |\nabla u|\,dx\,dy.
  \label{eq:TV}
\end{equation}

Let $u_*$ be extremal function for the following variational problem:
\begin{equation}
  u_* = \argmin_{u \in BV(\Omega)} J(u).
  \label{eq:ustar}
\end{equation}
The \autoref{eq:ustar} is called total variation regularization problem.

The norm and modulus of the gradient are metrical characteristic of a function
of two variables.
Continuous functions of two variables also have a set of topological
characteristics referred to as linear variations.
Kronrod%\cite{mak2}
introduced the notion of a regular component of the level set of a
continuous function of two variables.
The simplest topological characteristic in the linear variation theory is a
number of regular components for all level sets of a function.
Full information about linear variations of a function is contained in the
one-dimensional tree of a function of two variables.
In \cite{mak3,mak4,mak5}
was discussed the necessity of the using a linear variation in the
image restoration theory.

Let $\Phi_u(t)$ be the number of regular components of a level set $t$
for a continuous function.
The first Kronrod's linear variation is defined as
\begin{equation}
  V(u) = \int\limits_{-\infty}^{+\infty} \Phi_u(t)\,dt.
  \label{eq:KronrodVariation}
\end{equation}

Let $w$ be a binary discrete function $w=(w_{i,j})$, where $w_{i,j} \in
\{0,1\}$, for all pairs $(i,j)$. A subset of such pairs $(i, j)$ when
$w_{i,j}=1$ and all elements of the subset are connected by the 8-connectivity,
is called the connected component of the binary function $w$.
For a number $k \in \N$ and a discrete function $u$ we define the following
indicator function $\chi$:
\begin{equation}
  \chi_k (u_{i,j}) =
    \begin{cases}
      1, & u_{i,j} \ge k \\
      0, & u_{i,j} < k
    \end{cases}.
  \label{eq:chiIndicator}
\end{equation}

\begin{definition}
  The number $V_k(u_{i,j})$ of connected components for a level $k$, $k \in \N$
  of the discrete function $u$ is called the number of connected components of
  the binary discrete function $\chi_k(u_{i,j})$.
\end{definition}

\begin{definition}
  The linear variation $V(u_{i,j})$ of a discrete function  is defined as
  follows:
  \begin{equation}
    V(u_{i,j}) = \sum_{k=0}^{+\infty} V_k(u_{i,j})
    \label{eq:V}
  \end{equation}
\end{definition}

Let us compute the discrete gradient $\nabla u_{i,j}$ of $u$ at $(i,j)$ as
\begin{equation}
  \nabla u_{i,j} = (u_{i+1,j} - u_{i,j}, u_{i,j+1} - u_{i,j}).
  \label{eq:gradient}
\end{equation}

Suppose that if the pair $(i,j)$ is outside of the domain of the function $u$,
then $u_{i,j} = 0$.
The main problem of the computer implementation of the linear variation is 
computation of $V_k(u_{i,j})$ for given k.

\end{comment}

Continuous functions of two variables have a set of topological characteristics
referred to as linear variations. Kronrod\cite{mak2} introduced the notion of a
regular component of the level set of a continuous function of two variables.
The simplest topological characteristic in linear variation theory is the number
of regular components for all level sets of a function. In \cite{mak3,mak4,mak5}
was discussed the necessity of the using a linear variation in the image
restoration theory.

The problem of computing the linear variation in discrete case is equivalent 
to the problem of counting the number of connected components (CC) in a binary
image with 8-connectivity. This is usually done through CC labeling, and many
approaches for that are known\cite{hechao}. We came to the labeling problem
while searching the algorithm for linear variation computing. Of course, the
labeling problem also is well known in the digital image processing theory.

% Alekseev part

This paper presents an efficient way to convert the initial binary image with
8-connectivity into a condensed (non-binary) image with new connectivity that is
$2$ times smaller in each dimension than the initial image. Most approaches for
CC labeling are still valid for the new image. As time and memory consumptions
of conventional approaches usually depend linearly on the number of pixels,
running them on the condensed image can be more efficient than running on the
initial image.

It is worth noting that our method exploits an important property of
8-connectivity that does not hold for 4-connectivity, so it is not applicable
for the latter case.

\section{$2\times2$ condensation}

\texten{
We will use the fact that for the case of 8-connectivity all $4$ pixels of any
$2 \times 2$ square are adjacent to each other, and so are contained in the same
CC and should have same labels upon completion of the labeling algorithm. So
instead of labeling single pixels we can label $2 \times 2$ groups of pixels, or
``cells''.
}\textru{
Наблюдение: с 8-связностью в любом квадратике $2 \times 2$ любые два пикселя
находятся рядом. Значит если это пара объектных пикселей, то они находятся в
одной КС, и значит каждый такой квадратик пересекается не более чем с одной КС.
Поэтому вместо маркировки отдельных пикселей можно маркировать такие квадраты.
Назовём их клетками.
}

\texten{
Consider a $2N \times 2M$ binary image\footnote{
  For convenience in this paper we consider only input images with even size
  in both dimensions. If this is not the case, one can easily append a row or
  a column of background pixels to the image.
}
specified by a binary predicate $b(x, y)$, returning $1$ in case the pixel at
the given coordinates is black (object pixel), and $0$ otherwise (background
pixel). For convenience we use 0-indexation throughout the paper, so the topmost
row has row index $y=0$, and the leftmost column has column index $x=0$.
}\textru{
Рассмотрим бинарное изображение размера $2N \times 2M$\footnote{
  Для удобства мы рассматриваем только изображения с чётной высотой и шириной.
  В общем случае при необходимости можно добавить строку или столбец с фоновыми
  пикселями.
},
заданное предикатом $b(x, y)$, который возвращает $1$, если переданы координаты
объектного пикселя, и $0$ иначе.
Мы везде будем использовать $0$-индексацию: верхняя строка и левый столбец имеют
номер $0$.
}

\texten{
The condensed image will be of size $N \times M$. Each pixel of the condensed
image will unambiguously represent the configuration of the corresponding
$2\times 2$ cell of pixels of the initial binary image. So the condensed image
will contain pixels of $2^4 = 16$ colors. We suggest the function $c(x, y)$
denoting the color of a pixel of condensed image to be the following:
}\textru{
Сжатое изображение будет иметь размер $N \times M$. Каждый пиксель полученного
изображения должен будет однозначно задавать конфигурацию соответствующих
четырёх пикселей исходного изображения, поэтому он может быть одного из $2^4 =
16$ цветов. Предлагается такая функция $c(x,y)$, кодирующая четыре исходных
пикселя в цвет клетки:
}
\begin{align*}
  c(x, y) &= 2^0 b(2x, 2y)   + 2^1 b(2x+1, 2y) \\
          &+ {} 2^2 b(2x, 2y+1) + 2^3 b(2x+1,2y+1)
\end{align*}

\texten{
So the value of each pixel of the initial image is stored in the corresponding
bit in color of the corresponding pixel of the condensed image. Note that if
$c(x, y) = 0$ for some $(x, y)$, then there are no object pixels in this cell,
and no label should be assigned to this whole cell. The enumeration of pixels
inside one cell is shown on \autoref{fig:pixelsEnumeration}.
}\textru{
Таким образом значение каждого из исходных пикселей просто записано в
соответствующий бит в итоговом пикселе. Важно отметить, что если все четыре
исходных пикселя были фоновые, то и клетка будет иметь цвет 0. Нумерация
пикселей внутри клетки показана на \autoref{fig:pixelsEnumeration}.
}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node[cell] at (0.5, 0.5) {};
    \foreach \x/\y [count=\i from 0] in {0/1,1/1,0/0,1/0} {
      \node[whitebox] at (\x,\y) {\i};
    }
  \end{tikzpicture}
  \caption{Cell pixels enumeration}
  \label{fig:pixelsEnumeration}
\end{figure}

\texten{
We need also to define connectivity function for the condensed image.
We cannot actually just use 8-connectivity, as whether two cells are adjacent
cells of one CC depends not only on their coordinates, but also on their
configuration. See \autoref{fig:connectivity:examples} for examples.
}\textru{
Ещё нужно дать определение связности для сжатого изображения. Мы уже не можем
использовать 8-связность: теперь нельзя сказать, связаны ли две клетки, не
проанализировав их цвета. Смотри \autoref{fig:connectivity:examples}.
}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 1/1, 3/0, 3/1}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Nearby cells that are not directly connected.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 1/1, 2/0, 3/1}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Directly connected cells.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,4.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 0/1, 1/1, 4/0, 4/1, 5/0}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Cells that are not nearby in terms of 8-connectivity, cannot be
    directly connected.}
  \end{subfigure}
  \caption{Cells connectivity examples}
  \label{fig:connectivity:examples}.
\end{figure}

\texten{
However, for every variant of relative placement of two nearby cells there is a
mask of pixels of each cells that are important for direct connectivity of this
pair of cells. So two cells are considered directly connected, if they are
nearby in terms of 8-connectivity and each of them contains at least one object
pixel in the mask of important pixels corresponding to relative placement of the
cells. The masks are shown in \autoref{fig:connectivity:masks}.
}\textru{
Тем не менее, можно заметить, что анализировать можно не все биты цвета, а
только некоторую маску тех бит, которые важны для связности двух конкретных
клеток. Эти маски показаны на \autoref{fig:connectivity:masks}.
}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (0.5, 0.5) {};
      \node[graybox] at (0, 2) {2};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (0, 1) {0};
      \node[graybox] at (1, 1) {1};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 2.5) {};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 2) {2};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.25\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 0) {3};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 0) {2};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \caption{Masks of important pixels for all four cases of relative placement
    of two nearby cells. Those pixels are filled in gray, with their numbers
    within cell specified.}
  \label{fig:connectivity:masks}
\end{figure}

\texten{
The idea is that if we naturally encode mask pixels as 4-bit numbers in the
way similar to how we defined $c$, we can express the predicate denoting whether
two nearby cells
are directly connected efficiently with just two bitwise and one logical AND
operations.
As an example, \autoref{fig:connectivity:code} compares possible codes in \C
checking if the labels of the current pixel or cell and the one right on top of
it should be same for classic 8-connectivity and for the new connectivity. Note
that as bitwise operations are usually relatively cheap, we added only a little
complication compared to decreasing the number of pixels to process by the
factor of 4.
}\textru{
Идея в том, что с нашим естественным способом кодирования конфигурации клеток в
4-разрядных числах, предикат, определяющий связность двух соседных клеток,
может быть представлен всего двумя битовыми и одной логической операцией И.
В качестве примера, на \autoref{fig:connectivity:code} приведено сравнение
возможных фрагментов программы на языке \C\hspace{-0.8em}, проверяющих, связан
ли текущий пиксель (или клетка) с пикселем (или клеткой) над ним (ней).
Стоит отметить, что хотя конструкция усложнилась, битовые операции как правило
очень эффективны, и это усложнение, скорее всего, незначительно по сравнению с
уменьшением количества рассматриваемых позиций в 4 раза.
}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.475\linewidth}
    \centering
    \begin{lstlisting}
    if (b(x, y)) {
      if (b(x, y-1)) {
        // copy labels, or union label sets
      }
      // consider other directions
    }
    \end{lstlisting}
    \caption{Binary image with 8-connectivity.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.475\linewidth}
    \centering
    \begin{lstlisting}
    if (c(x, y)) {
      if ((c(x, y-1) & ((1<<2) | (1<<3)))
       && (c(x, y)   & ((1<<0) | (1<<1)))) {
        // copy labels, or union label sets
      }
      // consider other directions
    }
    \end{lstlisting}
    \caption{Condensed image with new connectivity.}
  \end{subfigure}
  \caption{Checking the nearby top pixel/cell.}
  \label{fig:connectivity:code}
\end{figure}

\section{Benchmarks}

In this section we present performance benchmarks of some popular methods for
counting CC both with and without the $2 \times 2$ cells optimization.

\subsection{Algorithms overview}

\texten{
As each of these methods, omitting DTSUF could be used with both condensed and
non-condensed images, for convenience we denote height and width of the image
on input of each algorithm as $H$ and $W$.
}\textru{
Так как все рассматриваемые методы, кроме DTSUF, можно применять как для
сжатого, так и для несжатого изображения, мы обозначим размеры входного
изображения каждого изображения за $H$ и $W$.
}

\texten{
Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
meaning the appropriate raster for non-condensed and condensed images. When we
mention neighbors of some pixel, we mean pixels that are directly connected to
it in terms of appropriate connectivity.
}\textru{
По аналогии, чтобы избежать повторения ``пиксель или клетка'' мы будет просто
использовать слово ``пиксель'', имея в виду растр входного изображения.
Под соседями пикселя будем понимать пиксели, прямо связанные с ним в терминах
соответствующей связности.
}

\subsubsection{DFS}

The first algorithm we chose for the benchmark is Depth First Search graph
traversal. Graph traversal algorithms are the general approach for CC labeling
in graphs, and
DFS is arguably the most simple.
Although not intended for image labeling and
suboptimal in most cases, it still has linear in the number of pixels worst-case
complexity in both time and memory.

Like most approaches, the algorithm has an outer loop over all pixels of the
image. Each time an object pixel (non-empty cell, in case it is actually a cell)
which is not labeled yet is encountered, DFS traversal is launched from that
pixel. The traversal labels all the object pixels of the same CC, launching
recursively from all unlabeled neighbors of current pixel.

\subsubsection{SUF}

\texten{
Another general approach for graph CC labeling which is more popular in image
labeling is using some kind of a data structure to maintain disjoint sets of
pixels that are considered to be in the same CC.
}\textru{
Другой, более популярный в этом контексте общий подход к подсчёту компонент
связности состоит в том, чтобы использовать структуру данных для поддержания
набора непересекающихся множеств пикселей, чем, по сути, КС и являются.
}

\texten{
The structure is commonly called Union-Find, as the two essential operations it
is supposed to support are Union, which is used to merge some two of the
disjoint sets in one, and Find, which is used to find the representative element
of the set some element belongs to, so that we can always tell if two elements
are in the same set. The structure can be implemented so that the amortized time
complexity for each operation is $O(\alpha(HW))$, where $\alpha$ is the inverse
Ackermann function, and $\alpha(HW) < 5$ for all remotely practical image
sizes\cite{CLRS}. % TODO links
}\textru{
Эту структуру принято называть Union-Find, так как две основные операции,
которые она поддерживает, это Union~--- объединение двух множеств в одно, и
Find~--- нахождение характерного представителя для множества, содержащего данный
элемент.
Этот представитель может быть использован для проверки принадлежности двух
элементов одному и тому же множеству.
Эта структура данных хорошо изучена и может быть реализована так, что
амортизированная сложность каждой операции~--- всего $O(\alpha(HW))$, где
$\alpha$~--- обратная функция Аккермана.
Для всех практически возможных размеров изображения эта функция не превосходит
5.
}

\texten{
The approach that is named SUF (Scan with Union-Find) in this paper follows.
}\textru{
Далее следует описание подхода, который мы называем SUF (Scan with Union-Find).
}

\texten{
Pixels are processed in the natural order: from top to bottom and from
left to right, so each row is fully processed before all rows after it.
So for each object pixel we encounter there are at most 4 nearby pixels
that were already processed, see \autoref{fig:suf:neighbors}.
}\textru{
Пиксели обрабатываются в естественном порядке сверху вниз слева направо, так что
каждая строка начинает обрабатываться, когда все предыдущие строки уже
обработаны. Таким образом, у очередного пикселя может быть до четырёх пикселей,
которые находятся рядом и уже были обработаны.
}

\texten{
At first each object pixel is considered to belong to its own CC, so right
before considering the neighbors we increment the CC counter. Moreover, it is
associated with its own disjoint set in the Union-Find structure. More
specifically, object pixel at position $x, y$ is associated with the set number
$xW + y$ to guarantee absence of collisions.
}\textru{
Каждый объектный пиксель сначала предполагается принадлежащим своей КС.
Поэтому прежде чем начинать рассматривать соседей, мы увеличиваем счётчик.
Более того того, каждый пиксель изначально ассоциирован со своим собственным
независимым множеством в структуре Union-Find, состоящем из одного элемента.
Конкретно пиксель в позиции $x, y$ заранее ассоциирован с элементом номер $xW +
y$, чтобы гарантировать отсутствие коллизий.
}

\texten{
The possible neighbors are processed in the order shown on
\autoref{fig:suf:neighbors}, and for each neighbor that is object pixel, the
Union operation is performed on the two sets containing the current and the
neighbor pixel. In case the Union operation was successful, which means that the
pixels had belonged previously to different sets and we have just found a way to
merge two CC, the CC counter is decremented.
}\textru{
Возможные соседи перебираются в порядке, показанном на
\autoref{fig:suf:neighbors}.
Для каждого объектного пикселя, который напрямую соседствует с текущим, мы
применяем операцию Union для множества, содержащих текущий пиксель и
множества, содержащего этого соседа.
Если эти множества не совпадали, то мы нашли место, где КС соединяются.
В этом случае уменьшаем счётчик на 1.
}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \foreach \x/\y [count=\i from 1] in {0/1,1/1,2/1,0/0} {
      \node[whitebox] at (\x,\y) {\i};
    }
    \node[graybox] at (1, 0) {};
    \node[whitebox] at (1, 0) {};
  \end{tikzpicture}
  \caption{Neighbor mask for SUF algorithm family.}
  \label{fig:suf:neighbors}
\end{figure}

\texten{
Note that this algorithm outputs not the matrix of final labels, but only the
number of CC. This is the place in which it differs from some common two-pass
approaches for CC labeling, as it is an algorithm not for labeling, but just for
counting CC. The latter is exactly what is needed for the aforementioned linear
variation problem.
}\textru{
Отметим, что этот алгоритм выдаёт не матрицу финальных меток, а только
количество компонент связности.
Именно в этом месте он отличается от некоторых известных подходов к CC labeling.
Для текущей задачи нам нужно именно считать компоненты связности.
}

\texten{
Also we have to mention that for performance reasons the Union-Find internal
structures are allocated ahead to have the size of at least $HW$ to guarantee
that each object can be easily associated with unique starting set given only
its coordinates. So memory consumption of SUF is linear in the number of pixels.
}\textru{
Также необходимо отметить, что для лучшей производительности служебные массивы
структуры Union-Find заранее выделяются размера как минимум $HW$, чтобы
обеспечить простоту соответствия между пикселями изображения и элементами в
системе множеств.
Таким образом, SUF использует линейное от числа пикселей количество памяти
$O(HW)$.
}

\subsubsection{SUF2}

\texten{
This approach is a modification of SUF designed to reduce memory consumption
and potential number of cache-misses generated by Union-Find operations.
}\textru{
Этот подход~--- модификация SUF, предложенная, чтобы сократить потребление
памяти и предположительно сильно уменьшить число кэш-промахов, происходящих
из-за произвольного доступа к большим массивам структуры Union-Find.
}

\texten{
The idea is to run through each row twice: first time to establish label
equivalence classes within row, and the second time to set unique equal label
for all pixels of same CC. This gives us the opportunity to maintain just $O(W)$
different labels at each moment, and so the Union-Find structure can be
allocated for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.
}\textru{
Идея в том, чтобы бежать по каждой строке дважды: один раз чтобы установить
классы эквивалентности меток в текущей и предыдущей строках, и второй раз, чтобы
расставить уникальные метки в пиксели каждого класса.
Это даёт нам возможность иметь не более $O(W)$ различных меток в каждый момент
времени, и поэтому Union-Find можно выделять размера всего $O(W)$, при том что в
SUF выделялось ровно $WH$ элементов.
}

\texten{
The exact maximum number of label equivalence classes between two adjacent rows
depends on whether the given image is condensed. For non-condensed images
$W=2M$, and maximum number of label equivalence classes in two adjacent rows is
$M=\frac{W}2$, as if two object pixels are both in adjacent rows and in adjacent
columns, then they are neighbors in terms of 8-connectivity.
}\textru{
Точная оценка на максимальное число классов эквивалентности меток в двух
соседних строках зависит от того, сжатое изображение поступило на вход, или нет.
Для несжатых изображений $W=2M$, и наибольшее число классов эквивалентности в
соседних строках равно $M=\frac{W}2$, так как если два объектных пикселя
находятся одновременно в соседних строках и соседних столбцах, то они соседи в
смысле 8-связности.
}

\texten{
On the other hand, for condensed images $W=M$, and maximum number of label
equivalence class in two adjacent rows of cells is $2W=2M$. See
\autoref{fig:suf2:extremal} for illustration on the extremal cases.
}\textru{
С другой стороны, для сжатых изображений $W=M$, и максимальное число классов
эквивалентности равно $2W=2M$. Смотри примеры на \autoref{fig:suf2:extremal}.
}

\begin{figure}
  \centering
  \begin{subfigure}{0.475\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0,1,...,9} \foreach \y in {0,1} {
        \node[whitebox,dashed,semithick,draw=gray] at (\x,\y) {};
      }
      \foreach \x/\y in {0/0,2/1,4/1,6/0,8/1} {
        \node[whitebox] at (\x,\y) {};
      }
    \end{tikzpicture}
    \caption{Non-condensed image, up to 5 different labels on 10 columns.}
  \end{subfigure}
  \quad
  \begin{subfigure}{0.475\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5,...,8.5} \foreach \y in {0.5,2.5} {
        \node[cell] at (\x,\y) {};
      }
      \foreach \x in {0,2,...,8} \foreach \y in {0,3} {
        \node[whitebox] at (\x,\y) {};
      }
    \end{tikzpicture}
    \caption{Condensed image, up to 10 different labels on 5 columns.}
  \end{subfigure}
  \caption{Extremal cases for SUF2.}
  \label{fig:suf2:extremal}
\end{figure}

\texten{
Our benchmarks show that, although SUF2 without condensation is generally more
efficient on noise images than SUF without condensation, the mentioned
consideration make the $2 \times 2$ optimization
of SUF2 less dramatic in means of performance increase factor.
Optimized SUF2 significantly outperforms optimized SUF only on
huge noise images with less than 60\% density, while falling behind on
small and medium-sized noise images.
}\textru{
Проведённые эксперименты показывают, что хотя SUF2 без сжатия в целом более
эффективен, чем SUF без сжатия, приведённая оценка делает относительное
ускорение SUF2 со сжатием не таким значительным.
}

\texten{
Note that, while SUF2 is intended for CC counting, its memory
consumption is only $O(W)$, which is significant for embedded systems and
other resource-tight applications.
}\textru{
Отметим, что SUF2 решает задачу подсчёта СК, используя всего $O(W)$
дополнительной памяти.
}

\subsubsection{DTSUF}

\texten{
This modification of SUF was implemented to compare the optimization we present
with another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo}, namely
decision tree. Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
applicable for condensed image, as it exploits similar observations on
8-connectivity.
}\textru{
Эта модификация SUF была добавлена чтобы сравнить нашу оптимизацию с
одной из оптимизаций, предложенных Wu, Otoo и Suzuki\cite{wuotoo}, а именно с
деревом принятия решения.
К сожалению, DTSUF (Decision Tree Scan with Union-Find) не применим для сжатого
изображения, так как использует свойства 8-связности, которыми наша новая
связность не обладает.
}

\texten{
As before, the goal is to reduce the number of pixel lookup and label copy
operations. This is achieved with careful consideration of only the necessary
neighbors, stopping once the set of one or two label operations that are needed
is determined. The decision tree for processing an object pixel is shown on
\autoref{fig:dtsuf}.
}\textru{
Как и раньше, цель состоит в том, чтобы уменьшить число запросов на содержание
пикселя и копирований меток.
Это достигается аккуратным рассмотрением соседей в оптимальном порядке и
остановкой рассмотрения, как только достаточная для выполнения всех необоходимых
действий информация была получена.
Смотри дерево на \autoref{fig:dtsuf}.
}

\begin{figure}
  \centering
  \begin{subfigure}{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \node[graybox] at (1,0) {};
      \foreach \x/\y/\a in {0/1/a,1/1/b,2/1/c,0/0/d,1/0/e} {
        \node[whitebox,font=\bf] at (\x,\y) {\a};
      }
    \end{tikzpicture}
    \caption{Neighbor enumeration.}
  \end{subfigure}
  \quad
  \begin{subfigure}{0.65\linewidth}
    \centering
    \begin{tikzpicture}[
      sibling distance=18em,
      lookup/.style = {shape=circle, draw, thick, align=center, font=\bf},
      action/.style = {shape=rectangle, rounded corners, draw, align=center, font=\scriptsize},
      result/.style = {font=\bf}
    ]
      \node[lookup] {b}
        child[level distance=3em,sibling distance=30em] {
          node[lookup] {c}
          child[level distance=5em, sibling distance=14em] {
            node[lookup] {a}
            child[level distance=6em, sibling distance=10em] {
              node[lookup] {d}
              child {
                node[action] {new label}
                edge from parent node[result,above] {$0$}
              }
              child {
                node[action] {union(e,d)}
                edge from parent node[result,above] {$1$}
              }
              edge from parent node[result,above] {$0$}
            }
            child {
              node[action] {union(e,a)}
              edge from parent node[result,above] {$1$}
            }
            edge from parent node[result,above] {$0$}
          }
          child[level distance=9em,sibling distance=24em] {
            node[lookup] {a}
            child[level distance= 6em, sibling distance=10em] {
              node[lookup] {d}
              child {
                node[action] {union(e,c)}
                edge from parent node[result,above] {$0$}
              }
              child {
                node[action] {union(e,c,d)}
                edge from parent node[result,above] {$1$}
              }
              edge from parent node[result,above] {$0$}
            }
            child[sibling distance=20em] {
              node[action] {union(e,c,a)}
              edge from parent node[result,above] {$1$}
            }
            edge from parent node[result,above] {$1$}
          }
          edge from parent node[result,above] {$0$}
        }
        child[level distance=2em] {
          node[action] {union(e,b)}
          edge from parent node[result,above] {$1$}
        };
    \end{tikzpicture}
    \caption{The decision tree.}
  \end{subfigure}
  \caption{DTSUF decision tree optimization.}
  \label{fig:dtsuf}
\end{figure}

\subsection{Condensation types}

\texten{
For memory-tight applications such as embedded systems it can be useful to
utilise the suggested condensation in a lazy way, so that without actually
constructing the condensed image we use a function (``view'') calculating color
of each cell with 4 lookups to the original binary image. This is probably most
interesting when combined with SUF2 algorithm, as it is a way to win some time
while avoiding memory footprint increase. However this lazy approach increases
the number of image lookups.
% So, according to our benchmark, this combination is efficient only for dense
% ($>30\%$) noise.
}\textru{
В некоторых случаях может быть более приемлемо использовать сжатие неявно, то
есть вместо построения 
}

So the benchmark charts show three levels of optimization applied to DFS, SUF
and SUF2: just method name in legend refers to method applied to non-condensed
image, `` on 2x2 view'' means lazy memory-less condensation and `` on 2x2 copy''
means that a temporary condensed image was constructed in memory. For the latter
case, the construction time was included into the overall measured time of run.

\subsection{Measurement}

For measurement purposes all algorithms and optimizations were implemented by us
in \CXX.

The program was compiled using \texttt{g++ 5.3.0-3} compiler with \texttt{-O3}
optimization flag. All tests were run on same machine running under Arch Linux
with standard \texttt{linux 4.4.1-2} kernel, utilising non-overclocked
\texttt{Intel(R) Core(TM) i5-2430M CPU @ 2.40GHz} (L3 cache size 3072K)
processor and non-overclocked 1333MHz 8GiB Kingston RAM.

The test images were random images, generated in the following manner. At first,
for given density $d$, the first $4dNM$ pixels of the $2N \times 2M$ image were
filled with object values, while rest filled with background values. Then all
the pixels were permuted using \texttt{std::shuffle} function from \CXX standard
library, which is an implementation of Fisher-Yates Shuffle algorithm.

All methods were run on same tests, which was achieved by supplying same seed to
the \texttt{std::mt19937} Mersenne Twister\cite{mt19937} random number engine
implementation from \CXX standard library before passing the engine to
\texttt{std::shuffle}.

Benchmarks showed that for random images, the run time depends non-trivially on
the value of density $d$, which is the ratio of object pixels in the image. On
the one hand, this is because all of the considered algorithms pass background
pixels without any processing, and somehow process the object pixels, so the
exact number of operations such as Union and Find should generally grow with
$d$. On the other hand, the greater the value of $d$, the less number of CC is
likely to be present in the image, which reduces the number of relatively
expensive Union operations.

So benchmarks were run for several values of $d$ to achieve a picture of how the
speedup depends on the number of object pixels in the image.

It is worth noting again that for ``2x2 copy'' optimization, the apparent time
overhead of constructing the extra condensed image  was included into the total
run time, since we measured the total time between calling the CC counting
subroutine, supplied with only the original image, and returning back from
it.

\subsection{Results}

The overall benchmark line plot of all algorithms and optimizations is
presented on \Cref{fig:overall}.

\begin{figure}
  \centering
  \includegraphics[height=0.45\textheight, keepaspectratio]{pics/overall.eps}
  \caption{Average CPU time per run on random images $1200 \times 1800$ px.}
  \label{fig:overall}
\end{figure}

To show more closely how efficient the optimizations actually appear to be, we
plotted speedup/density charts for the considered methods (\Cref{fig:normed}).
Here the speedup of algorithm A relative to algorithm B is defined as
$\frac{T_B}{T_A}$, where $T_A$ and $T_B$ are the total times used by the
corresponding algorithms to count CC in all the test images of same density.
The greater the speedup of optimized method relative to original method, the
more efficient the optimization is considered.

\newcommand{\inputnorm}[2]{
  \begin{subfigure}[t]{0.45\linewidth}
    \centering
    \includegraphics[width=\textwidth,
                     keepaspectratio]{pics/normed_#1.eps}
    \caption{Speedup of #2}
    \label{fig:normed:#1}
  \end{subfigure}
}

\begin{figure}
  \centering
  \inputnorm{DFS}{DFS optimizations}
  \inputnorm{SUF}{SUF optimizations}
  \inputnorm{SUF2}{SUF2 optimizations}
  \inputnorm{DTSUF}{SUF optimizations, relative to DTSUF}
  \caption{Speedup/density charts}
  \label{fig:normed}
\end{figure}

The speedup plots on \Cref{fig:normed:DFS,fig:normed:SUF,fig:normed:SUF2}
suggest that the ``2x2 copy'' optimization speeds up the considered methods by
factor greater than 1 for all tested densities. On the other hand, the
``2x2 view'' optimization is showed to be not time-efficient at all for some
densities, slowing down DFS and SUF2 if run on images with densities less than
38\% and 28\% respectively. Utilizing the latter optimization also does not
speed up SUF on images on density approaching 20\%, although it does not slow
down the algorithms either.

The worst-case and best-case speedups are shown on the \autoref{tab:speedup}.
Here we see that while ``2x2 view'' optimization is more memory-efficient than
``2x2 copy'', it can slow down the method in some situations, and generally it
does not give as much performance advantage.

We should also mention the chart on \autoref{fig:normed:DTSUF}, featuring
speedup of our SUF optimizations relative to DTSUF in order to compare our
optimizations to the optimization by Wu, Otoo and Suzuki. The plot clearly shows
that for all the densities under 80\% our ``SUF on 2x2 copy'' algorithm
outperforms the decision-tree based approach by the factor of at least $1.5$,
with speedup reaching $2.5$ for 2\% density.

\begin{table}
  \centering
  \begin{tabular}{l|c|c|c|c}
    Algorithm & worst speedup & worst speedup density & best speedup & best
    speedup density \\
    \hline
    \input{tables/speedup}
  \end{tabular}
  \caption{The worst and the worst speedups of the considered optimizations}
  \label{tab:speedup}
\end{table}

\section{Conclusion}

The problem of linear variation computing can be reduced to the problem of
counting connected components in a binary image with eight-connected
connectivity. This paper presents an efficient way to convert the initial binary
image with 8-connectivity into a condensed (non-binary) image with new
connectivity that is 2 times smaller in each dimension than the initial image.
Most approaches for connected components labeling are still valid for the new
image. As time and memory consumptions of conventional approaches usually depend
linearly on the number of pixels, running them on the condensed image can be up
to 4 times more efficient than running on the initial image. The method is
essentially a modification of some known raster algorithms for labeling that
groups pixels into $2 \times 2$ cells. A performance benchmark of the proposed
method applied to some approaches on noise images is provided. It is shown at
least up to 2 times more quick results could be received.

\section{Acknowledgements}

The work was supported by the Ministry of Education and Science of Russian
Federation (grant 2/1766/2014K).

\begin{thebibliography}{5}

%\bibitem {mak1}
% Rudin, L., Osher, S., Fatemi, E.:
% Nonlinear total variation based noised removal algorithms.
% Phys. D. 60, 259-268 (1992)

\bibitem {mak2}
Kronrod, A.:
On functions of two variables.
Uspehi Mat. Nauk 5, 1(35), 24-134(1950)

\bibitem {mak3}
Makovetskii, A., Kober, V.:
Image restoration based on topological properties of functions of two variables.
Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)

\bibitem {mak4}
Makovetskii, A., Kober, V.:
Modified gradient descent method for image restoration.
Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)

\bibitem {mak5}
Chochia, P., Milukova, O.:
Comparison of Two-Dimensional Variations in the Context of the Digital Image
Complexity Assessment.
Journal of Communications Technology and Electronics, 2015, 60, no. 12,  pp.
1432-1440

\bibitem {hechao}
He, L., Chao, Y., Suzuki, K., Wu, K.:
Fast connected-component labeling.
Pattern Recognition. v42, Pages 1977-1987. (2009)

%\bibitem {sterzh}
%Стержанов, М.:
%Методики выделения связных компонент в штриховых бинарных изображениях.
%Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. с. 169-174. (2010)

\bibitem {CLRS}
Cormen, H., Leiserson, E., Rivest, L., Stein, C.:
Introduction to Algorithms (Second ed.).
MIT Press (2001)

\bibitem {wuotoo}
Wu, K, Otoo, E., Suzuki, K.:
Optimizing two-pass connected-component labeling algorithms.
Pattern Anal. Appl. (2008)

% \bibitem {fisher:yates}
% Fisher, A., Yates, F.:
% Statistical tables for biological, agricultural and medical research (3rd ed.).
% London: Oliver \& Boyd. pp. 26–27. OCLC 14222135 (1948)

\bibitem{mt19937}
Matsumoto, M., Nishimura, T.:
Mersenne twister: a 623-dimensionally equidistributed uniform pseudo-random
number generator.
ACM Transactions on Modeling and Computer Simulation 8 (1): 3–30. (1998)

\end{thebibliography}

\end{document}